Trends of SDR in Thessaloniki

### 2023-07-03 Article_A.tex
Commit 2023-07-03 12:17

\title{Trends of SDR in Thessaloniki}
\author{true \and true}
\date{2023-06-30}

@@

The shortwave downward solar irradiance (SDR) at the Earth's surface plays a significant role, on Earths climate.
Changes of the SDR can be related to major changes on the Earth's energy budget, the mechanisms of climate change, and water and carbon cycle (Wild 2009).
Studies of SDR variability, have identified some distinct SDR trends on different regions of the world and for different time periods, using the term `brightening' for positive trends, and `dimming' for negative trends.
There are many cases on the long term records, showing a change of SDR trends direction, occurring roughly at the last decades of the 20th century.
A general dimming before 2000 and a brightening afterwards, was found on multiple station in China (Yang et al. 2021).
And a similar pattern was identified, with the breaking point around 1980, from a 71 years record of a Central Europe station (Wild et al. 2021) and for Brazil (Yamasoe et al. 2021).
Another approach, using AI aided spatial analysis on continent level, of numerous observational stations data, reach similar conclusions for the above regions and the global trend (Yuan, Leirvik, and Wild 2021).

There is a consensus, among researchers, that the major factors of the attenuation of SDR is the interaction with the atmospheric aerosols and the clouds.
Those interactions, among other factors, have been analysed with modeling (Li et al. 2016; Samset et al. 2018), showing the existence of feedback mechanisms between the two.
And similar finds, have been showed in observational data (Schwarz et al. 2020; Ohvril et al. 2009; Zerefos et al. 2009; Xia et al. 2007 and references therein).

Wild (2012) : \textasciitilde{}

Due to the variability of those factors and the interaction between them,

\ldots\ldots{} cloud formation occurrences and constitution nuclei. cite\ldots..

it is difficult to estimate the SDR trends on each site.

There is a also a need to evaluate the trends on different location

In this study, we examine the trends of SDR with ground-based measurements at Thessaloniki for the period 1993 - 2023, as derived from a CM-21 pyranometer.
We reevaluate and extend the dataset used by Bais et al. (2013), applying a different algorithm for the identification of clear-/cloud-sky instances, and we derive the radiation trends for this period under different sky conditions (all-sky, clear-sky and cloud-sky).

\ldots{}

Site location description and aerosols effect

\ldots{}

Our method to characterize the sky conditions and the definition of each sky condition has some subjectivity.
The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negatives identifications.

\hypertarget{data-and-methodology.}{%
\section{Data and methodology.}\label{data-and-methodology.}}

The SDR is equivalent to the whole sky Global radiation, also refered as Global Horizontal Irradiance (GHI), and was obtained with a horizontal leveled, CM-21 pyranometer.
SDR data spans the period of
1993-04-12 to 2023-03-31.
In some algorithms we used as auxiliary data, the direct beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer, with data availability from
2016-04-01 to 2023-03-31.

\ldots{}

\ldots{}

There are three distinct steps to the creation of this dataset:
a) the acquisition of radiation measurements from the sensors,
b) a radiation data quality check, and
c) the identification of ``clear sky'' conditions from the radiometric data
d) data aggregation and trend analysis.

For the acquisition of radiometric data,
the signal of the broadband instruments is sampled with a rate of \(1 \text{Hz}\). The mean and standard deviation values are recorded for every minute.
The measurements are corrected for the zero offset of the instrument signal.
As reference of the ``dark signal'' is used the Sun elevation angle bellow \(-10^\circ\), for a period of \(3 \text{h}\).
The signal is converted to radiation flux, using a ramped value of the instrument sensitivity, derived from the eight laboratory calibrations of the instrument, during the study period.

A manual screening is performed, to remove inconsistencies and erroneous recordings, that can occur randomly or systematically, during the long continuous operation of the instrument.
The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi~(C. Long and Shi 2008 ; C. Long and Shi 2006).
Thus, problematic recordings have been excluded from further processing.
Although it is not possible to detect all the bad data, the large number of data and the aggregation scheme we use, can provide us with accurate radiometric measurements, for the scope of this study.

\hypertarget{clear-sky-identification}{%
\subsection{Clear sky identification}\label{clear-sky-identification}}

@@

As global radiation clear sky reference we are using the Haurwitz's model, adjusted for our site with a factor of 0.965 (Eq. \ref{eq:ahau}).
The selection of a clear sky reference model, was based on SDR observation from the period 2016 -- 2021.
Where, after an iterative optimization of eight simple models (Daneshyar--Paltridge--Proctor, Kasten--Czeplak, Haurwitz, Berger--Duffie, Adnot--Bourges--Campana--Gicquel, Robledo-Soler, Kasten and Ineichen-Perez) with different factors.
We found, that Haurwitz's model, adjusted with a factor of 0.965 has the lower root mean squared error (RMSE).
The tried models are described by Reno, Hansen, and Stein (2012) and tested by Reno and Hansen (2016).
The iterative optimization method, for the selection of the reference is discussed by C. N. Long and Ackerman (2000) and Reno and Hansen (2016).

@@

In order to estimate the effect of the clouds on the SDR we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition.
The all-sky conditions, containing all the valid measurements.
The clear-sky conditions data, where we have inferred that the sky was almost clear of clouds, and the remainder part the data as cloudy sky conditions data (cloud-sky).
To identify the clear-sky conditions we used the method proposed by C. N. Long and Ackerman (2000) and by Reno and Hansen (2016), that was adapted and configured for the site.
We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, due to the used method of characterization.
As a result, the details of the definition are site specific, it relies on a combination of threshold of comparisons with ideal actinometric models and statistics on different signal behavior.

\hypertarget{data-and-data-selection.}{%
\subsection{Data and data selection.}\label{data-and-data-selection.}}

Due to a significant measurement uncertainty near, the horizon, we have to exclude
all measurements with SZA greater than \(85^\circ\).
Moreover, due to some obstructions around the site (hills and buildings), we excluded data with Azimuth angle between
\(35^\circ\) and \(120^\circ\) with SZA greater than \(80^\circ\).
On the latter instances, Sun is systematically, not visible by the instrument's location.
To make the measurements comparable throughout the dataset, we adjusted all 1-minute radiometric values to the mean Sun - Earth distance.
Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at \(1 \text{au}\), in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
The TSI data we use, comes from NOAA (Coddington et al. 2005).
As a result, we can present all radiation data as a fraction of Sun's TSI.

@@

\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions.}\label{fig:trendALL}
\end{figure}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2048}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1928}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2169}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3855}}@{}}
\caption{\label{tab:trendtable}Trends of daily means by sky conditions.
}\tabularnewline
\toprule
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend statistical signif. {[}\%{]}
\end{minipage} \\
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend statistical signif. {[}\%{]}
\end{minipage} \\
\midrule
\endhead
0.3996 & 8.665e-23 & All sky cond. & 100 \\
0.3773 & 7.075e-18 & Clear sky cond. & 100 \\
0.3594 & 3.318e-09 & Cloudy cond. & 100 \\
\bottomrule
\end{longtable}

@@

\caption{Trends by season for all condition. Displaying monthly means of daily means.}\label{fig:seasonalALL}
\end{figure}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Reno2012}{}}%
Reno, Matthew J., Clifford W. Hansen, and Joshua S. Stein. 2012. {``Global Horizontal Irradiance Clear Sky Models: Implementation and Analysis.''} SAND2012-2389, 1039404. \url{https://doi.org/gq5npv}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Wild2012}{}}%
---------. 2012. {``Enlightening Global Dimming and Brightening.''} \emph{Bulletin of the American Meteorological Society} 93 (1): 27--37. \url{https://doi.org/b3w2h4}.

>>>>

\title{Trends of SDR in Thessaloniki}
\author{true \and true}
\date{2023-07-03}

@@

The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on it's climate.
Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle (Wild 2009).
Can also affect, solar and agricultural production, and living organisms.
Studies of SDR variability, have identified some distinct SDR trends on different regions of the world and for different time periods.
The term `brightening' is generally used to describe periods of positive SDR trend, and `dimming' for negative trend.
There are many cases on the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
On multiple station in China, a dimming period was reported until about 2000, followed by a brightening period (Yang et al. 2021).
A similar pattern was identified, with the breaking point around 1980, for stations in Central Europe (Wild et al. 2021) and Brazil (Yamasoe et al. 2021).
Also, on global scale, an AI aided continental level spatial analysis, with data from multiple station, reach similar conclusions for the above regions and for the global trend (Yuan, Leirvik, and Wild 2021).

There is a consensus, among researchers, that major factors of SDR attenuation is the interaction with atmospheric aerosols and clouds.
Those interactions, among other factors, have been analysed with models (Li et al. 2016; Samset et al. 2018), showing the existence of feedback mechanisms between the two.
Similar finds, have been showed in observational data (Schwarz et al. 2020; Ohvril et al. 2009; Zerefos et al. 2009; Xia et al. 2007 and references therein).

Due to the variability of the phenomenon, and its contributing factors, there is a need to further investigate and monitor SDR, in different sites, to estimate its magnitude, and its relation to the local conditions.
In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki, Greece for the period 1993 to 2023, as derived from a CM-21 pyranometer.
We reevaluated and extended the dataset used by Bais et al. (2013), applying a different algorithm for the identification of clear-/cloud-sky instances (Reno and Hansen 2016; Reno, Hansen, and Stein 2012a), and we derive the radiation trends for the period 1993 to 2023, under different sky conditions (all-sky, clear-sky and cloud-sky).

\hypertarget{observational-data-and-methodology.}{%
\section{Observational data and methodology.}\label{observational-data-and-methodology.}}

The SDR is equivalent to the whole sky Global radiation (or Global Horizontal Irradiance, GHI), and was obtained with a horizontal leveled, CM-21 pyranometer, from
1993-04-12
to
2023-03-31.
Also, we have measurements of direct beam radiation (DNI) obtained by a collocated CHP-1 pyrheliometer, from
2016-04-01
to
2023-03-31.
Although, we performed a similar analysis to the DNI as for the SDR, the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data.
Moreover, the DNI was used as auxiliary data, with the clear sky identification algorithm (CSid), to aid with the selection of the appropriate thresholds.
We have to note also, that despite the capability of the CSid to use the DNI as an characterization parameter, we haven't use it, to avoid any selection bias due to unequal availability of measurements.

The SDR data were measured with a Kipp \& Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki (\(40^\circ38'\)N, \(22^\circ57'\)E, 80m a.s.l.).
The monitoring site is located near the city centre.
During the study period the CM-21 pyranometer has been independently calibrated four times.
There are three distinct steps to the creation of this dataset:
a) the acquisition of radiation measurements from the sensors,
b) a radiation data quality check, and
c) the identification of ``clear sky'' conditions from the radiometric data
d) data aggregation and trend analysis.

For the acquisition of radiometric data,
the signal of the broadband instruments is sampled with a rate of \(1 \text{Hz}\). The mean and its standard deviation are recorded for every minute.
The measurements are corrected for the zero offset of the instrument signal.
As reference of the ``dark signal'' is used the Sun elevation angle bellow \(-10^\circ\), for a period of \(3 \text{h}\).
The signal is converted to radiation flux, using a ramped value of the instrument sensitivity, derived from the eight laboratory calibrations of the instrument, during the study period.

A manual screening was performed, to remove inconsistencies and erroneous recordings, that can occur randomly or systematically, during the long continuous operation of the instrument.
The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi~(2008, 2006).
Thus, problematic recordings have been excluded from further processing.
Although it is not possible to detect all the bad data, the large number of data and the aggregation scheme we use, can provide us with accurate radiometric measurements, for the scope of this study.

In order to estimate the effect of the clouds on the SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition.
The all-sky conditions, containing all the valid measurements.
The clear-sky conditions data, correspond to data where we have inferred that the sky was almost clear of clouds, and the remainder part the data was characterized with cloud sky conditions (cloud-sky).
To identify the clear-sky conditions we used the method proposed by C. N. Long and Ackerman (2000) and by Reno and Hansen (2016), that was adapted and configured for the site.

cite iterative process

@@

We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, due to the used method of characterization.
As a result, the details of the definition are site specific, it relies on a combination of threshold of comparisons with ideal actinometric models and statistics on different signal behavior.

The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negatives identifications.

As global radiation clear sky reference we are using the Haurwitz's model, adjusted for our site with a factor of 0.965 (Eq. \ref{eq:ahau}).
The selection of a clear sky reference model, was based on SDR observation from the period 2016 -- 2021.
Where, after an iterative optimization of eight simple models (Daneshyar--Paltridge--Proctor, Kasten--Czeplak, Haurwitz, Berger--Duffie, Adnot--Bourges--Campana--Gicquel, Robledo-Soler, Kasten and Ineichen-Perez) with different factors.
We found, that Haurwitz's model, adjusted with a factor of 0.965 has the lower root mean squared error (RMSE).
The tried models are described by Reno, Hansen, and Stein (2012b) and tested by Reno and Hansen (2016).
The iterative optimization method, for the selection of the reference is discussed by C. N. Long and Ackerman (2000) and Reno and Hansen (2016).

@@

Due to a significant measurement uncertainty near, the horizon, we have to exclude
all measurements with SZA greater than \(85^\circ\).
Moreover, due to some obstructions around the site (hills and buildings), we excluded data with Azimuth angle between
\(35^\circ\) and \(120^\circ\) with SZA greater than \(80^\circ\).
On the latter instances, Sun is systematically, not visible by the instrument's location.
To make the measurements comparable throughout the dataset, we adjusted all 1-minute radiometric values to the mean Sun - Earth distance.
Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at \(1 \text{au}\), in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
The TSI data we use, comes from NOAA (Coddington et al. 2005), it overlaps with our data, and is interpolated to match with the time step of the data.
As a result, we can present all radiation data as a fraction of Sun's TSI.

@@

\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023  `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`. The black line shows the long term trend for all-sky conditions.}(\#fig:trendALL)
\end{figure}

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2361}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2222}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2500}}@{}}
\caption{\label{tab:trendtable}Trends of daily SDR means, for different sky conditions. All cases have a statistical significance above \(99.9\%\)
\normalsize}\tabularnewline
\toprule
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} \\
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} \\
\midrule
\endhead
0.3996 & 8.665e-23 & All sky cond. \\
0.3773 & 7.075e-18 & Clear sky cond. \\
0.3594 & 3.318e-09 & Cloudy cond. \\
\bottomrule
\end{longtable}

@@

\caption{Trends by season for all condition. Displaying monthly means of daily means.}(\#fig:seasonalALL)
\end{figure}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Reno2012a}{}}%
Reno, Matthew J., Clifford W. Hansen, and Joshua S. Stein. 2012a. {``{Global Horizontal Irradiance Clear Sky Models: Implementation and Analysis}.''} \emph{SANDIA REPORT Sand2012-2389 Unlimited Release Printed March 2012}, March, 1--66.

\leavevmode\vadjust pre{\hypertarget{ref-Reno2012}{}}%
---------. 2012b. {``Global Horizontal Irradiance Clear Sky Models: Implementation and Analysis.''} SAND2012-2389, 1039404. \url{https://doi.org/gq5npv}.

@@
